### 2020-03-02 doc/icfp20/intro.tex

[NOTE: I've commented out the discussion on explicit / implicit
handling of rules, to make the flow smoother.]

% Full dependent types allow expression of predicate logic, and it is
% tempting to take a standard textbook on music theory such as
% \citet{piston-harmony} or \citet{aldwell2018harmony} and 
% formalize it in type theory. However this is difficult because most
% ``rules'' in music theory are not absolute, but rather suggestions of
% varying degrees (themselves vaguely specified) of importance. Since
% ultimately the resulting created music is more important than the
% theory, an alternative is to attempt to indirectly deduce the rules
% from music samples, using machine learning~\citep{huang-cp} or
% some other statistical methods. Unfortunately the set of deduced
% rules could sometimes be insufficient for generating good music.
% As an example, \citet{roberts-cp} develop the Bach Doodle
% harmonization tool using Bach chorals as training data, but still
% observe high frequency of parallel 5th and 8th that are extremely
% rare in Bach's music.

@@

Our long-term goal is to formalize a large amount of music theory.
As a first step we start with the modest task of expressing a small,
relatively strict rule set known as species
counterpoint~\cite{fux-cp}, intended for combining interdependent
melody lines to produce a pleasant-sounding result.
Roughly speaking, what we do is to write a custom musical type
checker in Agda.
Compared to the previous work by \citet{cong-cp}, where rules are
expressed directly as Agda types, our approach makes it easier to
describe fine-grained rules, produce readable error messages, and
add or drop rules depending on the circumstances.

[NOTE: I've squeezed the description of our new counterpoint
implementation.]

>>>>

% However this is difficult because most
% ``rules'' in music theory are not absolute, but rather suggestions of
% varying degrees (themselves vaguely specified) of importance. Since
% ultimately the resulting created music is more important than the
% theory, an alternative is to attempt to indirectly deduce the rules
% from music samples, using machine learning~\citep{huang-cp} or
% some other statistical methods. Unfortunately the set of deduced
% rules could sometimes be insufficient for generating good music.
% As an example, \citet{roberts-cp} develop the Bach Doodle
% harmonization tool using Bach chorals as training data, but still
% observe high frequency of parallel 5th and 8th that are extremely
% rare in Bach's music.

@@

Full dependent types allow expression of predicate logic, and it is
tempting to take a standard textbook on music theory such as
\citet{piston-harmony} or \citet{aldwell2018harmony} and formalize it
in type theory.  As a first step we start with the modest task of
expressing a small, relatively strict rule set known as species
counterpoint~\cite{fux-cp}, intended for combining interdependent
melody lines to produce a pleasant-sounding result.  Roughly speaking,
what we do is to write a custom musical type checker in Agda.
Compared to the previous work by \citet{cong-cp}, where rules are
expressed directly as Agda types, our approach makes it easier to
describe fine-grained rules, produce readable error messages, and add
or drop rules depending on the circumstances.

% [NOTE: I've squeezed the description of our new counterpoint
% implementation.]
